### 2018-01-18 Manuscript.Rmd
Added a notebook to collect all the figures and stats we will use for the manuscript.

---
title: "Exposure Consequences - Statistics and Figures"
output:
  pdf_document: default
  html_notebook: default
---
Here I collect the 'right' statistics and figures for the exposure consequences project.

All the scripts doing the statistics and figures are in separate R files, but here we will only see / do the ones that we think should go to the manuscript.

First we load those other scripts:

```{r}
source('shared.R')
source('nocursor.R')
source('localization.R')
source('relateLocalizationNoCursors.R')
```

# Reach Aftereffects

Here is a plot:

```{r fig.width=10, fig.height=4}
plotReachAftereffects()
```

Messages of the figure:

Panel A: there are substantial and persisting reach aftereffects
Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)