### 2018-06-23 ManuscriptMaterials.Rmd
small changes

---
title: "Exposure Consequences - Statistics and Figures"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
---

```{r setup, cache=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(comment='')
```

# Overview

This document collects statistics and figures to be used in the manuscript on the exposure consequences project.

## Figures

These are the planned figures. Figure one will not be included here, and can in fact simply be re-used from the previous manuscript.

1. Setup & Task (goal: visualize the experiment)
    + A: Setup (side view)
    + B: Rotated training (top view)
    + C: No-cursor reaches (top view)
    + D: Localization (top view)
2. Reach Aftereffects (goal: show motor changes are there, are robust and compare to classic)
    + A: Exposure reach aftereffects (initial & later)
    + B: Classic & Exposure reach aftereffects (initial only? or all?)
3. Localization
    + A: Exposure, active and passive (goal: small/no difference between active and passive)
    + B: Classic, active and passive (goal: larger difference between active and passive, different generalization?)
    + C: Online localization?

Maybe:

4. Correlation
    + A: Exposure, active and passive localization predicting reach aftereffects
    + B: Classic, active and passive localization predicting reach aftereffects

## Tables

We expand the table used in the previous manuscript to include this paradigm as well.

1. Task order & trial numbers (goal: clarify task order and show similarity between classic and exposure experiment - so probably after Fig 1)

## Source scripts

The scripts doing the statistics and creating figures are in separate R files. They allow many options, but here we will only see / do the ones that we think should go to the manuscript.

First we load those other scripts:

```{r loadsources}
source('shared.R') # functions used everywhere
source('nocursor.R') # functions for no-cursor reach data
source('localization.R') # functions for localization data
source('relateLocalizationNoCursors.R') # functions that correlate the two kinds of data
```

## Topics

In the manuscript we'll first show that there indeed are motor changes after exposure training (with the no-cursor data) and then compare that with the classic. Second is the localization: a) localization shifts after exposure b) the shifts are (not) different from those after classic c) the effects are (not) different across the workspace d) this pattern is (not) different from that after classic. A third topic might be if localization shifts can predict no-cursor changes, which we'll call the "correlation"  topic. Here we'll also follow this order:

1. Reach aftereffects
2. Localization
3. Correlations

## Analyses

### ANOVA or LME

Since in the localization tasks there is missing data at some points (usually at the extremes) we can't use ANOVAs directly on the data. Imputation is going to be very hard because the missing data is systematically missing. Hence, we are using Linear Mixed Effects models as an alternative, but this presents a new problem. Getting the usual F-values and p-values is not so straightforward. Initially we decided to use the $\chi^2$ approach that we get from the `Anova` function from the `car` package. However, a paper from Steven G. Luke (2017, _Evaluating significance in linear mixed-effects models in R_) suggests using Sattertwaite approximation. So all functions running an LME now take an argument that tells them to use the Satterthwaite method, but the default is chi-squared. It turns out that it doesn't matter much, so it's a good sanity check. For consistency across all the analyses, we set the value of that argument is set here:

```{r}
LMEmethod <- 'Satterthwaite'
# options: Satterthwaite, chi-squared
```

### Peak location of generalization

Since participants trained with a single target, we could look at generalization curves. These appear to be shifted more counter clockwise in most cases, so we want to quantify this. But, for most of the curves, and especially the ones that seems to be shifted, we don't have data on the full curve on one side of the peak. Or maybe it doesn't even include the peak itself. That makes it hard to fit a normal function to the data. Instead, we assume that the minimal change in the data, occurs far beyond the boundaries of the measured workspace, make the data cumulative and scale back to a range from 0 to 1. This allows fitting a _cumulative_ normal function where the 50% point corresponds to the peak of the generalization curve. This method is more stable, and actually works on all the datasets used (but we didn't test online localization from the classic experiment). However, the results should still be considered carefully.

# Reach Aftereffects

Here is a plot, probably Figure 2:

```{r fig2_reachaftereffect, fig.width=7.5, fig.height=2.5}
plotReachAftereffects()
```

Messages of the figure:

- Panel A: there are substantial and persisting reach aftereffects
- Panel B: that are somewhat lower than those in classic (but not that much? but only for some targets?)

These claims require analyses.

## Is there an effect of exposure training on reach aftereffects?

First, we show that no-cursor reaches change direction immediately after exposure training (iteration 1):

```{r RAE_exposure,include=TRUE}
exposureNoCursorChange(LMEmethod=LMEmethod)
```

So this shows that whether or not the feedback was rotated mattered for the reach deviations in the exposure training group, i.e.: there are reach aftereffects.

## Do the motor changes persist during localization?

The second analysis is to show that the reach aftereffects aren't all that different for first and later iterations of the rotated no-cursor tasks. (The rotated data is always compared with _all_ aligned iterations of the task.)

```{r RAE_exp_iterations}
exposureAftereffectsPersistent(LMEmethod=LMEmethod)
```

So there is no effect of iteration here, nor does it interact with target, which means the reach aftereffects are the same for the first iteration and the average across the other four iterations. We now also have enough power to see an effect of target angle: the generalization curve is not flat.

## Are reach aftereffects comparable between classic and exposure training?

Finally, we want to see if there is a difference between the reach aftereffects observed after exposure training and those after classic training.

```{r RAE_exp_cla}
exposureClassicReachAftereffects(LMEmethod=LMEmethod)
```

I'd have expected there to be some difference between the two training types, perhaps only in interaction with target, but that's not the case apparently.

The absence of an interaction between training and target could be interpreted to mean that the generalization of reach aftereffects is the same after both kinds of training. Here are the confidence intervals for the peaks of both:

```{r RAE_peaks}
getPeakConfidenceInterval('classic', part='all', CIs=c(.95))
getPeakConfidenceInterval('exposure', part='all', CIs=c(.95))
```

The 95% confidence interval for the classically trained group includes the 45 degree target (where the visual feedback was during training) but not the 75 degree target, whereas the 95% confidence interval for the exposure training group includes the 75 degree target (where the hand really was during training) but not the 45 degree target. Something might be going on -- but the two confidence intervals are not different from each other, which is in line with the LME results. We can't conclude much, except that we have no evidence for different generalization curves.

To sum up:

1. Rotated exposure training evokes substantial shifts in no-cursor reach direction across the workspace,
2. these changes persist over time, and
3. these changes are comparable to the ones we found after classic training, and
4. there is some generalization curve across the workspace, but
5. we have no evidence that this is different from reach aftereffect generalization after classic training.

# Localization

Here is a plot, probably figure 3:

```{r fig3_localization, fig.width=7.5, fig.height=2.5}
plotLocalization(classicOnline=TRUE)
```

It seems that just like after classic training, there is a change in localization responses following exposure training. There might still be a difference between active and passive localization after exposure training - contrary to our expectations - but it could be smaller than the difference after classic training. This will need to be tested.

## Is there a localization shift after exposure training?

We first want to see if there is any overall effect of aligned versus rotated training in the exposure group (i.e.: are there any localization shifts at all?), with movement type and hand angle as fixed effects, and particiant as random effect.

```{r loc_exp}
exposureLocalization(remove15=TRUE, LMEmethod=LMEmethod)
```

Wether or not the feedback was rotated makes a difference according to two terms in the model. In other words: exposure training caused localization responses to shift (systematically), as we could already guess by looking at the figure. This means we can look at the difference between repsonses after rotated and after aligned feedback: the training induced shift in localization. The shifts are also different, given different hand angles, which means there is some form of generalization pattern.

## Effects of movement type and hand angle in exposure localization

Let's first see if there is a difference between active and passive localization after exposure training.

```{r loc_exp_movtype}
exposureLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
```

There is no difference between passive and active localization after exposure training, even with double the number of trials as in classic delayed.

## Comparing localization after exposure and classic training

Let's see what we get in a model with all data (localization differences between rotated and aligned though) from both groups:

```{r loc_ombnibus}
groupLocalization(model='full', remove15=TRUE, LMEmethod=LMEmethod)
```

In the exposure group there is no effect of movement type, but we see one here. This could be because the effect of movement type is much larger in the classic group. This would predict an interaction between group and movement type, but that's not the case (p=.133).

What if we don't care about hand angle:

```{r loc_groups_movementtype}
groupLocalization(model='movementtype', remove15=TRUE, LMEmethod=LMEmethod)
```

Still no effect...

Let's see if the classic group still shows an effect of movement type (with the current preprocessing procedures):

```{r loc_cla}
classicLocalizationShift(remove15=TRUE, LMEmethod=LMEmethod)
```

So there is a fairly strong effect of movement type in the classic group. Together with the absence of an effect of movement type in the exposure group, but no interaction between group and movement type if we combine both datasets.

Also note the absence of an effect of hand angle in the classic group here -- probably because we excluded the 15 degree target, but it seems that after classic training the generalization curve for delayed localization is much flatter.

To sum up our findings on localization:

1. Rotated exposure training leads to changes in localization responses.

2. In the exposure group, we find no difference between localization responses with and without efference copies.
3. In the classic group, the localization responses informed by efference copies, are shifted more than those with only proprioception (as found before).
4. This predicts an interaction between movement type and group in the combined analysis, but we don't see that.

5. The shift in localization is different across the workspace in the exposure group. I.e., there is a generalization _curve_.
6. This effect is not present in the classic group, so that generalization can be said to be flat; there is no _curve_. **NOTE:** When we do include the 15 degree target, there is an effect of hand angle though.

## Localization peaks

Here are the confidence intervals for the peaks of the generalization curves for _active_ localization:

```{r localization_peaks}
getPeakLocConfidenceInterval(group='classic',CIs=c(.95), movementtype='active', LRpart='all')
getPeakLocConfidenceInterval(group='exposure',CIs=c(.95), movementtype='active', LRpart='all')
```

These results mimic those of the reach aftereffects: the median values are different, but they each fall in the others confidence interval. However, here the LME showed that there was an interaction between training type and hand angle, which suggests that the generalization curves are different. If we want to look at this better, we need to get better estimates of these curves in every participant, by having a wider range of points and more data per point, but probably also larger rotations.

# Correlation

In this section we look at how recalibrated proprioception (hand localization) correlates with reach aftereffects.

This plot shows regression lines with localization shifts as predictor of reach aftereffects, potentially Figure 4:

```{r scattergrams, fig.width=7.5, fig.height=2.5}
correlations <- correlateNoCursorsLocalization()
```

It looks like for exposure, active localization is the better predictor of reach aftereffects, whereas in classic it is passive localization. But are any of them real? The function also returns the correlation tests and linear regressions for all four datasets.

## Pearons Correlations

As one possibility, we can look at the correlations:

```{r pearson_rhos}
for (name in names(correlations)) {
  cat(name)
  print(correlations[[name]][['cortest']])
}
```

Only one is significant: active localization predicting reach aftereffects in the exposure group.

## Linear regression models

And those linear regression models?

```{r linreg, eval=FALSE, include=FALSE}
# this code won't run or show up in output, just included here if you want to play with it:
for (name in names(correlations)) {
  cat(name)
  print(correlations[[name]][['linearm']])
}
```

[code/results hidden] They get the same p-value, so that seems fine.

## Multiple regression models

We've decided to use all the no-cursor blocks for these regression and correlations, so perhaps we shouldn't include the one below where I only use the initial block. The idea behind using only the initial block is that that initial block matches the time the localization is done -- relative to the last training.

Here I do step-wise multiple regression, allowing active and passive localization shift as predictor of reach aftereffects. If I do this on all the no-cursor block data, the models converge on using the active localization for the exposure group, and passive localization for the classic group.

If we look at only the initial no-cursor block, the story changes. It uses both active and passive to predict reach aftereffects for the exposure group, but converges on using only an intercept term for the classic group. This might mean that in te classic group there is no relationship between reach aftereffects and localization, or this might mean that in the exposure group we have sufficient amount of measurements to detect the relationship, as we have twice the number of localization trials. One way to rule this out is to use only the first half of the localization data from the exposure group.

Here is what the step-wise multiple regressions look like with all available data:

```{r step_exposure}
multipleRegressionLocalization(group='exposure',NCpart='initial')
```

No terms are dropped an the adjusted R squared of the full model is low, but reasonable. _Note that the coefficient for the passive localization is positive._

For classic training, it looks like this:

```{r step_classic}
multipleRegressionLocalization(group='classic',NCpart='initial')
```

The step-wise multiple regression algorithm thinks an intercept-only model is best. Now is this because localization after classic training has little to no information on reach aftereffects, or is this because we have double the data in the exposure group? We can test this partially by throwing away the second half of the data for the exposure group:

```{r step_exposure_firsthalf}
multipleRegressionLocalization(group='exposure',NCpart='initial', LRpart='first')
```

Now the passive localization shifts are thrown out by the algorithm and the active localization shifts don't contribute much either. This means that perhaps previous papers didn't find any relationships between recalibrated proprioception and reach aftereffects because not enough data was collected to overcome the noisiness of the measurements?

It also makes me wonder if reach aftereffects in classic can be better predicted by the online localization responses:

```{r step_online}
multipleRegressionLocalization(group='online',NCpart='initial')
```

But no.

# Conclusions

What can we conclude? Here is an initial list of what we might conclude.

1. Exposure training leads to robust reach aftereffects that are similar to those after classic training, although they seem to have a more pronouced generalization curve. This means that evoking propceptive recalibration with visual-proprioceptive discrepancies leads to motor changes. In other words: proprioception informs motor control.

2. Exposure training leads to shifts in localization, that are not very different between active and passive localization. Since both active and passive movements access recalibrated proprioception, while active localization also accesses updated predictions, this indicates that hand localization is makes very good use of proprioception.

    1. Additionally, since predictions should not be updated in exposure training, active localization could have resulted in a _lower_ shift in localization if hand location were estimated by combining recalibrated proprioception and updated prediction in a maximum likelihood way. If anything however, active localization is slightly larger (non-significant) than passive localization. This indicates a non-Bayesian process of combining predictions and perceptions -- in this case.

    2. After classic training (2016 paper) there was an effect of movement type, which indicates that with a training regime that allows updates of predicted sensory consequences, these updates do indeed also affect hand localization.

    3. Within the workspace however, there is no interaction between movement type and training group (but there is a 'trend', p=.09, if hand angle is used as a random effect, in all analyses the 15 degree data is omitted as Denise said). Perhaps there is a small update of predicted proprioceptive consequences, that we can't parcel out. A more computational approach would be needed to see if this can explain the pattern of results here.

3. Recalibrated proprioception can somewhat predict reach aftereffects, but only after **exposure** training and for _active localization_. This indicates that in some circumstances, recalibrated proprioception might drive motor changes. However, it is possible that the active component of the localization is crucial here. Perhaps exposure training leads to some updates in predicted sensory consequences after all -- be they visual or proprioceptive. These would also factor into the reach aftereffects.

    1. There is also a trend towards significance for _passive localization_ to predict reach aftereffects after **classic training**. Perhaps active localization is more similar to no-cursor reaches after exposure training, as they both use recalibrated proprioception as well as updated predictions of sensory consequences -- which could mostly be proprioceptive consequences after exposure training. This does not explain the trend towards significance for predicting classic reach aftereffects with passive localization. A step-wise multiple regression favours an intercept model for classic reach aftereffects, hence none of the localization data was deemed informative. The same algorithm arrived at a model using active localization to predict exposure reach aftereffects. The intercept model for classic reach aftereffects could signify that not too much should be read into the trend towards significance in the correlation. On the other hand, the exposure group had twice as much data, so that the estimated recalibration would be more reliable within each participant. Perhaps with higher quality estimates of proprioceptive recalibration those would be able to predict part of the reach aftereffects even after classic training.

```{r create_svgs, cache=FALSE, include=FALSE}
# This chunk creates SVG files but shouldn't show up in the PDF / HTML 
plotReachAftereffects(generateSVG=TRUE)
plotLocalization(classicOnline=TRUE, generateSVG=TRUE)
correlations <- correlateNoCursorsLocalization(generateSVG=TRUE)
```